Section Rings of $\bq$-divisors on minimal rational surfaces

### 2017-11-24 surfaces-write-up/surface-paper.tex
finished going through comments and added response to referee

\subsection{Notation}
\todo{move this notation subsection to its own section, after the introduction}
Throughout, we work over a fixed algebraically closed field $\bk$.
%Since generator and relation degrees are preserved under base change to the algebraic closure, 
%as cohomology commutes with flat base change,
%assuming that $\bk$ is algebraically closed is no restriction
%on generality and the results hold equally well over an arbitrary field.
\begin{rem}
	Because cohomology commutes with flat base change, in particular with field extensions,
the dimensions of the graded pieces of the section ring will be preserved under base change from $\bk$ to $\overline \bk$. Therefore generators
and relations are preserved under arbitrary base field extension, and hence their minimal degrees are preserve.
Because of this, there is not harm in assuming $\bk =\overline \bk$. The bounds we give below hold equally well
over arbitrary fields, but for simplicity we state and prove them over algebraically closed fields.

We note that canonical divisors of root stacks may not behave well with respect to base change of purely inseparable extensions,
so base change along purely inseparable extensions may cause problems if one considers 
some section ring associated to the canonical divisor.
But, for the purposes of this article, we fix a divisor on a scheme over some base field, and the structure of that particular section
ring is unchanged upon base change to the algebraic closure.
\end{rem}

We now define some notation used throughout the paper.
We use $m \in \bz_{\geq 0}$ to index the dimension of a given projective
space $\bp^m$ and the type of the Hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes_\bz \bq.
\end{align*}\noindent
where $n \in \bz_{\geq 0}$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. In the case $X = \bp^m$, define the degree of $D$ by $\deg D :=
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. The floor of a $\bq$-divisor $D$ is the divisor
$\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor D_i$. Let 
$R_D := \bigoplus_{d \geq 0} H^0(\bp^m, \lfloor dD \rfloor)$
denote the section ring associated to the $\bq$-divisor $D$. 
We often alternatively write
$R_D := \bigoplus_{d \geq 0} u^d H^0(\bp^m, \lfloor dD \rfloor)$, where $u$ is a dummy
variable to keep track of the degree.
We use $R_D$ as notation for $R(X,D)$ when $X$ is understood from context.
If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. 
If $D \in \di X \otimes_\bz \bq$
is an arbitrary divisor, we denote $h^0(X, D) := \dim_\bk H^0(X,D)$.

@@

In Section \ref{sec:proj} we prove the following two results bounding
the degree of generators and relations of
section rings on $\bp^m$. The first applies
to effective $\bq$-divisors and the second applies to arbitrary
$\bq$-divisors.

@@

Then $R_D$ is generated in degrees at most 
$\sum_{i = 0}^n \ell_i a_i$
with relations generated in degrees at most
\[
	\max \left(2 \sum_{i=0}^n \ell_i a_i, \frac{\max_{0 \le i \le n}
	(	\bida_i)}{\deg(D)} + \sum_{i=0}^n \ell_i a_i \right).
\]
\end{restatable}

@@

Note that the assumption given in Theorem~\ref{thm:proj-generators-relations}
that $\left\{ f_0, \ldots, f_n \right\}$ contains a basis for $H^0(\bp^m, \sco_{\bp^m}(1))$ is not a serious assumption:
If $D$ is an arbitrary divisor, then $D + 0 \cdot V(x_0) + \cdots + 0 \cdot V(x_m)$ is a divisor satisfying this assumption. Plugging this divisor into the statement
of Theorem~\ref{thm:proj-generators-relations}, we obtain the following bounds for an arbitrary divisor:
\begin{cor}
	\label{cor:proj-generators-relations}
	Let $D = \sum_{i = 0}^n \alpha_i D_i \in \di \bp^m \otimes_\bz \bq$ be a divisor of positive degree
with $\alpha_i = \frac{c_i}{k_i}\in \bq$ in reduced form. Write
$\ell := \lcm_{0 \leq j \leq n}(k_j)$, $\ell_i := \lcm_{0\le j \le n, j \ne i}(k_j)$ and $a_i := \deg D_i$. Let $\bp^m
\cong \proj \bk[x_0, \ldots, x_m]$ and let
$f_i \in \bk[x_0, \ldots, x_m]$ such that $D_i = V(f_i)$.

Then $R_D$ is generated in degrees at most 
$\ell(m+1) + \sum_{i = 0}^n \ell_i a_i$
with relations generated in degrees at most
\[
	\max \left(2 \left(\ell(m+1)+\sum_{i=0}^n \ell_i a_i \right), \frac{\max_{0 \le i \le n}
	(	\bida_i)}{\deg(D)} + \ell(m+1) + \sum_{i=0}^n \ell_i a_i \right).
\]
\end{cor}

@@

Then $R_D$ is generated in degrees at most
\[
	\lcm_{1\le i\le n}(k_i)\cdot \left(\sum_{1\le i\le j\le n} \bida_i \bidb_i\right) 
\]
\noindent
with relations generated in degrees at most 
\[
	2 \cdot \lcm_{1\le i\le n}(k_i) \cdot \left( \sum_{1\le i\le j\le n} \bida_i \bidb_i\right) 
\]
\end{thm}
\begin{rem}
Theorem~\ref{thm:hirz-generators-relations} is restated with a more precise
bound in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}
\begin{rem}
	As in Theorem~\ref{thm:proj-generators-relations}, the condition that $D$ contain independent polynomials
	in $u,v$ and $w,z$ is easily removed by replacing $D$ with $D' :=D +0 \cdot u + 0 \cdot v + 0 \cdot w + 0 \cdot z$,
	and computing the resulting bound for $D'$.
	One can either take the simpler bound (but slightly weaker) bound from Theorem~\ref{thm:hirz-generators-relations}
	or the more precise bound given in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}

@@

@@

\noindent
has a presentation over $R_{D'}$ consisting of the $\sum_{i = 0}^{r}
{{m + c_i} \choose {c_i}}$ generators $F_i^{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $0 \leq i \leq r$ and $\vec{v} \in \bz_{\geq 0}^{m + 1}$
with $\deg \vec v = c_i$. Furthermore, the ideal of
relations $I$ is generated by the following two classes of elements.
\begin{enumerate}
	\item For each $(i, j)$ with $j \geq i + 2$ and each $\vec{v} \in \mss_i,
		\vec{w} \in \mss_j$, there is either a relation of the form
		\begin{align*}
			G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
			- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})
			^{g_{\vec{y}}} \in I,
		\end{align*}
		with $i < h_{i, j} < j$, or there is a relation of the form
		\begin{align*}
			G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
			- \left( \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})^{g_{\vec{y}}}
			\right) \left( \prod_{ \vec{z} \in
			\mss_{h_{i, j} + 1}} (F_{h_{i, j} + 1}^{\vec{z}})^{g'_{\vec{z}}} \right) \in I,
		\end{align*}
		with $i < h_{i, j} < h_{i, j} + 1 < j$.
	\item For each $(i, j)$ with
		$j = i$ or $j = i + 1$ and each $\vec{v} \in \mss_i, \vec{w} \in
		\mss_j$ with $\vec{v} \not\prec \vec{w}$ (see Definition
		~\ref{defn:vec-order}) there is a relation of the form
		\begin{align*}
			&L_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
			- F_i^{\vec{y}} F_j^{\vec{z}} \in I
		\end{align*}

@@

\noindent
\todo{should we rule out other stupid relations, like those given when $w < v$?}
where $\vec{y}$ and $\vec{z}$ are the unique
vectors in $\mss_i$ and $\mss_j$, respectively, such that $\vec{y}
+ \vec{z} = \vec{v} + \vec{w}$ and $\vec{y} \prec \vec{z}$.
\end{thm}

@@

Before giving bounds for arbitrary divisors on projective space in Subsection~\ref{ssec:arbitrary-projective},
we give a detailed example of the generators and relations for a
section ring associated to an effective divisor on projective space.
\begin{example}
	\label{example:}
	In this example, we work out generators and relations 
	for the section ring $R(\bp^2, D)$ with
	\begin{align*}
		D := \frac{1}{6}V(x^2 + y^2 + z^2) + \frac{2}{5}V(x).
	\end{align*}
	We loosely follow the
	algorithm
	described in the course of the 
	proof of Theorem~\ref{thm:proj-effective-intro} and use notation from the statement of Theorem~\ref{thm:proj-effective-intro}.

@@

The main aim of this section is to prove Theorem
~\ref{thm:proj-generators-relations} bounding the number of generators and
relations of $R_D$. 
Having justified the necessity of adding ghost divisors, we proceed to bound the
number of generators and relations of arbitrary $\bq$-divisors in 
projective space. 
In PRoposition~\ref{prop:pm-span-and-basis}, we record a general proposition describing
a basis for $H^0(\bp^m, dD)$.
We bound the generators in Lemma ~\ref{lem:proj-generators}, and we use Lemmas ~\ref{lem:composite-map}, ~\ref{lem:bound-ker-chi}, and ~\ref{lem:proj-relations-psi} to bound the 
degree of relations in the proof of Theorem ~\ref{thm:proj-generators-relations}. Proposition ~\ref{prop:cone-generation}, Lemma ~\ref{lem:composite-map}, and Lemma ~\ref{lem:bound-ker-chi} are quite 
general and will also be used in Section ~\ref{sec:hirz} to bound the 
degree of generators and relations on Hirzebruch surfaces.
However, before moving on to the proof of ~\ref{thm:proj-generators-relations}, we justify the importance of assumption
~\eqref{eqn:pm-basis-assumption} with several illustrative examples.
%riga
\sssec*{The importance of ghost divisors}
\label{sssec:ghost-divisors}

@@

\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
(1) and (2) lie in $H^0(\mathbb{P}^m,dD)$, and are also necessary to lie in
$H^0(\mathbb{P}^m,dD)$.
To complete the proof, it
suffices to check functions satisfying conditions (1)-(3) form
a basis of $H^0(\bp^m,dD)$. There are $\binom{m+ \deg \lfloor dD \rfloor }{m}$
functions satisfying conditions (1)-(3). However we know
$h^0(\bp^m,dD) = \binom{m+ \deg \lfloor dD \rfloor }{m},$ so it suffices to show
that those monomials satisfying conditions (1)-(3) are
independent. 
To see why these are independent, observe that monomials satisfying (1)-(3) are all of the form
$f_0^{c_0} \cdots f_m^{c_m} \cdot g$ for the fixed monomial
$g = f_{m+1}^{-\lfloor d\alpha_{m+1} \rfloor} \cdots f_n^{-\lfloor d\alpha_{m+1} \rfloor}$.
For any fixed $N \in \bz$, the set 
\begin{align*}
 \left\{ f_0^{c_0} \cdots f_m^{c_m} : (c_0, \ldots, c_m) \in \bz^{m+1}, \sum_{i=0}^m c_i = N \right\}	
\end{align*}
forms an independent set over $\bk$.
Therefore, multiplying the monomials in the above set with the $g$ also forms an independent set, and these are precisely the monomials satisfying (1)-(3) 
with $N = -\deg g$.
\end{proof}

>>>>

Throughout, for ease of notation,
we work over a fixed algebraically closed field $\bk$.
This is nonessential, see \autoref{rem:algebraically-closed}

@@

After briefly stating our notation in Section~\ref{section:notation},
in Section~\ref{sec:proj}, we prove the following two results bounding
the degree of generators and relations of
section rings on $\bp^m$. The first applies
to effective $\bq$-divisors and the second applies to arbitrary
$\bq$-divisors.

@@

Then $R_D$ is generated in degrees at most 
$\omega := \sum_{i = 0}^n \ell_i a_i$
with relations generated in degrees at most
\[
	\max \left(2 \omega, \frac{\max_{0 \le i \le n}
	(	\bida_i)}{\deg(D)} + \omega \right).
\]
\end{restatable}

@@

Note that the assumption given in Theorem~\ref{thm:proj-generators-relations}
that $\left\{ f_0, \ldots, f_n \right\}$ contains a basis for $H^0(\bp^m, \sco_{\bp^m}(1))$ is not a serious assumption:
If $D$ is an arbitrary divisor, then $D + 0 \cdot V(x_0) + \cdots + 0 \cdot V(x_m)$ is a divisor satisfying this assumption. Plugging this divisor into the statement
of Theorem~\ref{thm:proj-generators-relations}, we obtain the following bounds for an arbitrary divisor:
\begin{cor}
	\label{cor:proj-generators-relations}
	Let $D = \sum_{i = 0}^n \alpha_i D_i \in \di \bp^m \otimes_\bz \bq$ 
with $\alpha_i = \frac{c_i}{k_i}\in \bq$ in reduced form. Write
$\ell := \lcm_{0 \leq j \leq n}(k_j)$, $\ell_i := \lcm_{0\le j \le n, j \ne i}(k_j)$ and $a_i := \deg D_i$. Let $\bp^m
\cong \proj \bk[x_0, \ldots, x_m]$ and let
$f_i \in \bk[x_0, \ldots, x_m]$ such that $D_i = V(f_i)$.

Then $R_D$ is generated in degrees at most 
$\omega' := \ell(m+1) + \sum_{i = 0}^n \ell_i a_i$
with relations generated in degrees at most
\[
	\max \left(2 \omega', \frac{\max_{0 \le i \le n}
	(	\bida_i)}{\deg(D)} + \omega' \right).
\]
\end{cor}

@@

Then $R_D$ is generated in degrees at most
\[
	\rho' := \lcm_{1\le i\le n}(k_i)\cdot \left(\sum_{1\le i\le j\le n} \bida_i \bidb_i\right) 
\]
\noindent
with relations generated in degrees at most 
$2 \rho'$.
%\[
%	2 \cdot \lcm_{1\le i\le n}(k_i) \cdot \left( \sum_{1\le i\le j\le n} \bida_i \bidb_i\right) 
%\]
\end{thm}
\begin{rem}
Theorem~\ref{thm:hirz-generators-relations} is restated with a more precise
bound in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}
\begin{rem}
	\label{rem:hirz-generators-relations-hypothesis}
	As in Theorem~\ref{thm:proj-generators-relations}, the condition that $D$ contain independent polynomials
	in $u,v$ and $w,z$ is easily removed by replacing $D$ with $D' :=D +0 \cdot u + 0 \cdot v + 0 \cdot w + 0 \cdot z$,
	and computing the resulting bound for $D'$.
	One can either take the simpler bound (but slightly weaker) bound from Theorem~\ref{thm:hirz-generators-relations}
	or the more precise bound given in Theorem~\ref{thm:hirz-generators-relations-full}.
\end{rem}

@@

\section{Notation}
\label{section:notation}
\invisiblesubsection
%Since generator and relation degrees are preserved under base change to the algebraic closure, 
%as cohomology commutes with flat base change,
%assuming that $\bk$ is algebraically closed is no restriction
%on generality and the results hold equally well over an arbitrary field.

In this section, we now collect various notation, used throughout the paper.
Throughout, for ease of notation,
we work over a fixed algebraically closed field $\bk$.
Our results hold equally well over arbitrary fields, as we now explain.

\begin{rem}
	\label{rem:algebraically-closed}
	Because cohomology commutes with flat base change, in particular with field extensions,
the dimensions of the graded pieces of the section ring will be preserved under base change from $\bk$ to $\overline \bk$. Therefore generators
and relations are preserved under arbitrary base field extension, and hence their minimal degrees are preserve.
Because of this, there is not harm in assuming $\bk =\overline \bk$. The bounds we give below hold equally well
over arbitrary fields, but for simplicity we state and prove them over algebraically closed fields.

We note that if $L/\bk$ is an inseparable extension,
and $X$ is a scheme over $\bk$ than the base change of the canonical divisor
$(K_X)_L$ may be different then the canonical divisor of the base change
$K_{X_L}$.
Therefore, the canonical ring may not be preserved under base change
along inseparable extensions.
But, for the purposes of this article, we fix a divisor on a scheme over some base field, and the structure of that particular section
ring is unchanged upon base change to the algebraic closure.
\end{rem}

We next define some notation used throughout the paper.
We use $m \in \bz_{\geq 0}$ to index the dimension of a given projective
space $\bp^m$ and the type of the Hirzebruch surface $\hirz_m$.
Let $D$ be a (Weil) $\bq$-divisor of the form
\begin{align*}
	D = \sum_{i=1}^{n}\alpha_i D_i \in \di X \otimes_\bz \bq.
\end{align*}\noindent
where $n \in \bz_{\geq 0}$ indexes the number of irreducible divisors in 
the above expansion of $D$, $\alpha_i \in \bq$, and $D_i \in \di X$ is an integral
codimension 1 closed subscheme of $X$. When it is convenient to do
so, we shall sometimes start the indexing at $0$, so that $i$ runs
from $0$ to $n$. In the case $X = \bp^m$, define the degree of $D$ by $\deg D :=
\sum_{i=1}^{n}\alpha_i \cdot \deg D_i$. The floor of a $\bq$-divisor $D$ is the divisor
$\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor D_i$. Let 
$R_D := \bigoplus_{d \geq 0} H^0(X, \lfloor dD \rfloor)$
denote the section ring associated to the $\bq$-divisor $D$. 
We often alternatively write
$R_D := \bigoplus_{d \geq 0} u^d H^0(X, \lfloor dD \rfloor)$, where $u$ is a dummy
variable to keep track of the degree.
We use $R_D$ as notation for $R(X,D)$ when $X$ is understood from context.
If $R$
is a graded ring, we denote the $d$th graded component of $R$ by $R_d$. If $r$ is a rational number, we let
$\fr(r) := r - \lfloor r
\rfloor$ denote the fractional part of $r$. 
If $D \in \di X \otimes_\bz \bq$
is an arbitrary divisor, we denote $h^0(X, D) := \dim_\bk H^0(X,D)$.

@@

\noindent
has a presentation over $R_{D'}$ consisting of the $\sum_{i = 0}^{r}
{{m + c_i} \choose {c_i}}$ generators $F_i^{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $0 \leq i \leq r$ and $\vec{v} \in \bz_{\geq 0}^{m + 1}$
with $\deg \vec v = c_i$. Furthermore, the ideal of
relations $I$ is generated by the following two classes of elements.
\begin{enumerate}
	\item For each $(i, j)$ with $j \geq i + 2$ and each $\vec{v} \in \mss_i,
		\vec{w} \in \mss_j$, there is either a relation of the form
		\begin{align*}
			G_{i, j}^{\vec{v}, \vec{w}} := F_i^{\vec{v}} F_j^{\vec{w}}
			- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})
			^{g_{\vec{y}}} \in I,
		\end{align*}
		with $i < h_{i, j} < j$, or there is a relation of the form
		\begin{align*}
			G_{i, j}^{\vec{v}, \vec{w}} := F_i^{\vec{v}} F_j^{\vec{w}}
			- \left( \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})^{g_{\vec{y}}}
			\right) \left( \prod_{ \vec{z} \in
			\mss_{h_{i, j} + 1}} (F_{h_{i, j} + 1}^{\vec{z}})^{g'_{\vec{z}}} \right) \in I,
		\end{align*}
		with $i < h_{i, j} < h_{i, j} + 1 < j$.
	\item For each $(i, j)$ with
		$j = i$ or $j = i + 1$ and each $\vec{v} \in \mss_i, \vec{w} \in
		\mss_j$ with $\vec{v} \not\prec \vec{w}$ (see Definition
		~\ref{defn:vec-order}) there is a relation of the form
		\begin{align*}
			&L_{i, j}^{\vec{v}, \vec{w}} := F_i^{\vec{v}} F_j^{\vec{w}}
			- F_i^{\vec{y}} F_j^{\vec{z}} \in I
		\end{align*}

@@

\noindent
where $\vec{y}$ and $\vec{z}$ are the unique
vectors in $\mss_i$ and $\mss_j$, respectively, such that $\vec{y}
+ \vec{z} = \vec{v} + \vec{w}$ and $\vec{y} \prec \vec{z}$.
\end{thm}

@@

Before giving bounds for arbitrary divisors on projective space in Subsection~\ref{ssec:arbitrary-projective},
we give a detailed example of the generators and relations for a
section ring associated to an effective divisor on projective space.
\begin{example}
	\label{example:effective-example}
	In this example, we work out generators and relations 
	for the section ring $R(\bp^2, D)$ with
	\begin{align*}
		D := \frac{1}{6}V(x^2 + y^2 + z^2) + \frac{2}{5}V(x).
	\end{align*}
	We loosely follow the
	algorithm
	described in the course of the 
	proof of Theorem~\ref{thm:proj-effective-intro} and use notation from the statement of Theorem~\ref{thm:proj-effective-intro}.

@@

The main aim of this section is to prove Theorem
~\ref{thm:proj-generators-relations} bounding the number of generators and
relations of $R_D$. 
Having justified the necessity of adding ghost divisors, we proceed to bound the
number of generators and relations of arbitrary $\bq$-divisors in 
projective space. 
In PRoposition~\ref{prop:pm-span-and-basis}, we record a general proposition describing
a basis for $H^0(\bp^m, dD)$.
We bound the generators in Lemma ~\ref{lem:proj-generators}, and we use Lemmas ~\ref{lem:composite-map}, ~\ref{lem:bound-ker-chi}, and ~\ref{lem:proj-relations-psi} to bound the 
degree of relations in the proof of Theorem ~\ref{thm:proj-generators-relations}. Proposition ~\ref{prop:cone-generation}, Lemma ~\ref{lem:composite-map}, and Lemma ~\ref{lem:bound-ker-chi} are quite 
general and will also be used in Section ~\ref{sec:hirz} to bound the 
degree of generators and relations on Hirzebruch surfaces.
However, before moving on to the proof of ~\ref{thm:proj-generators-relations}, we justify the importance of assumption
~\eqref{eqn:pm-basis-assumption} with several illustrative examples.

\sssec*{The importance of ghost divisors}
\label{sssec:ghost-divisors}

@@

\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
(1) and (2) lie in $H^0(\mathbb{P}^m,dD)$.
Conditions (1) and (2) are also necessary for some monomial in the $f_i$ to lie in
$H^0(\mathbb{P}^m,dD)$.
Since the monomials in the $f_i$ span $R_D$, it follows the monomials satisfying
$(1)$ and $(2)$ span $(R_D)_d = H^0(\bp^m, dD)$.
To complete the proof, it
suffices to check functions satisfying conditions (1)-(3) form
a basis of $H^0(\bp^m,dD)$. There are $\binom{m+ \deg \lfloor dD \rfloor }{m}$
functions satisfying conditions (1)-(3). However we know
$h^0(\bp^m,dD) = \binom{m+ \deg \lfloor dD \rfloor }{m},$ so it suffices to show
that those monomials satisfying conditions (1)-(3) are
independent. 
To see why these are independent, observe that monomials satisfying (1)-(3) are all of the form
$f_0^{c_0} \cdots f_m^{c_m} \cdot g$ for the fixed monomial
$g = f_{m+1}^{-\lfloor d\alpha_{m+1} \rfloor} \cdots f_n^{-\lfloor d\alpha_{m+1} \rfloor}$.
For any fixed $N \in \bz$, the set 
\begin{align*}
 \left\{ f_0^{c_0} \cdots f_m^{c_m} : (c_0, \ldots, c_m) \in \bz^{m+1}, \sum_{i=0}^m c_i = N \right\}	
\end{align*}
forms an independent set over $\bk$.
Therefore, multiplying the monomials in the above set with the $g$ also forms an independent set, and these are precisely the monomials satisfying (1)-(3) 
with $N = -\deg g$.
\end{proof}
